Analysis for Project

### 2021-11-14 analysis/analysis.Rmd
Code design matrix
I have created a design matrix from the analytical dataset that is used for the actual models. It allows me to easily mean center everything and then set defaults for the predict command.

```{r}
#try a model
model0 <- glm(below_exp_grade~as.factor(year)+current_grade+as.factor(state)+metro, 
              data=acs, family=binomial)
model1 <- update(model0, .~.+race)
model2 <- update(model1, .~.+foreign_born+foreign_born_mother+foreign_born_father)
model3 <- update(model2, .~.+degree_mother+degree_father)
model4 <- update(model3, .~.+sqrt(family_income/10000)+own_home+parents_married)

knitreg(list(model1, model2, model3, model4))
```

```{r}
#run separate models by grade groups
model4a <- update(model4, data=subset(acs, as.numeric(current_grade)<=6))
model4b <- update(model4, data=subset(acs, as.numeric(current_grade)>6 &
                                        as.numeric(current_grade)<=9))
model4c <- update(model4, data=subset(acs, as.numeric(current_grade)>=10))

@@

>>>>

```{r design-matrix}

metro <- model.matrix(~metro, acs)[,-1]
colnames(metro) <- c("met_non_metro","met_central_city","met_non_central_city",
                     "met_city_indeter")
#mean center
metro <- t(t(metro)-apply(metro, 2, mean))
metro <- as.data.frame(metro)

years <- model.matrix(~as.factor(year), acs)[,-1]
colnames(years) <- paste("y", 2011:2019, sep="")
#mean center
years <- t(t(years)-apply(years, 2, mean))
years <- as.data.frame(years)

states <- model.matrix(~as.factor(state), acs)[,-1]
colnames(states) <- paste("s", 1:50, sep="")
#mean center
states <- t(t(states)-apply(states, 2, mean))
states <- as.data.frame(states)

grades <- model.matrix(~current_grade, acs)[,-1]
#mean center
grades <- t(t(grades)-apply(grades, 2, mean))
grades <- as.data.frame(grades)

degree <- model.matrix(~degree_mother+degree_father, acs)[,-1]
#mean center
degree <- t(t(degree)-apply(degree, 2, mean))
degree <- as.data.frame(degree)

design_matrix <- cbind(states, years, grades, metro, degree)

design_matrix$foreign_born <- acs$foreign_born-mean(acs$foreign_born)
design_matrix$foreign_born_mother <- acs$foreign_born_mother-mean(acs$foreign_born_mother)
design_matrix$foreign_born_father <- acs$foreign_born_father-mean(acs$foreign_born_father)

design_matrix$income_trans <- sqrt(acs$family_income/10000)
design_matrix$income_trans <- design_matrix$income_trans-mean(design_matrix$income_trans)
design_matrix$own_home <- acs$own_home-mean(acs$own_home)
design_matrix$parents_married <- acs$parents_married-mean(acs$parents_married)

design_matrix$race <- acs$race
design_matrix$below_exp_grade <- acs$below_exp_grade
```

```{r}
#start with raw race differences
model0 <- glm(below_exp_grade~race, data=design_matrix, family=binomial)
model1 <- update(model0, .~.+y2011+y2012+y2013+y2014+y2015+y2016+y2017+y2018+y2019+
                current_grade1st+current_grade2nd+current_grade3rd+current_grade4th+
                current_grade5th+current_grade6th+current_grade7th+current_grade8th+
                current_grade9th+current_grade10th+current_grade11th+current_grade10th+
                current_grade11th+current_grade12th+
                met_non_metro+met_central_city+met_non_central_city+met_city_indeter+
                s1+s2+s3+s4+s5+s6+s7+s8+s9+s10+s11+s12+s13+s14+s15+s16+s17+s18+s19+
                s20+s21+s22+s23+s24+s25+s26+s27+s28+s29+s30+s31+s32+s33+s34+s35+s36+
                s37+s38+s39+s40+s41+s42+s43+s44+s45+s46+s47+s48+s49+s50)
model2 <- update(model1, .~.+foreign_born+foreign_born_mother+foreign_born_father)
model3 <- update(model2, .~.+degree_motherHS+degree_motherAA+degree_motherBA+
                   degree_motherG+degree_fatherHS+degree_fatherAA+
                   degree_fatherBA+degree_fatherG)
model4 <- update(model3, .~.+income_trans+own_home+parents_married)

knitreg(list(model0, model1, model2, model3, model4))
```

```{r eval=FALSE}
#run separate models by grade groups
model4a <- update(model4, data=subset(acs, as.numeric(current_grade)<=6))
model4b <- update(model4, data=subset(acs, as.numeric(current_grade)>6 &
                                        as.numeric(current_grade)<=9))
model4c <- update(model4, data=subset(acs, as.numeric(current_grade)>=10))

@@

```{r}
#dateset for predict
df_pred <- data.frame(race=levels(acs$race),
                      y2011=0, y2012=0, y2013=0, y2014=0, y2015=0, y2016=0,
                      y2017=0, y2018=0, y2019=0,
                      current_grade1st=0, current_grade2nd=0, current_grade3rd=0,
                      current_grade4th=0, current_grade5th=0, current_grade6th=0,
                      current_grade7th=0, current_grade8th=0, current_grade9th=0,
                      current_grade10th=0, current_grade11th=0, current_grade12th=0,
                      met_non_metro=0, met_central_city=0, met_non_central_city=0,
                      met_city_indeter=0,
                      s1=0, s2=0, s3=0, s4=0, s5=0, s6=0, s7=0, s8=0, s9=0, s10=0,
                      s11=0, s12=0, s13=0, s14=0, s15=0, s16=0, s17=0, s18=0, s19=0,
                      s20=0, s21=0, s22=0, s23=0, s24=0, s25=0, s26=0, s27=0, s28=0,
                      s29=0, s30=0, s31=0, s32=0, s33=0, s34=0, s35=0, s36=0, s37=0,
                      s38=0, s39=0, s40=0, s41=0, s42=0, s43=0, s44=0, s45=0, s46=0,
                      s47=0, s48=0, s49=0, s50=0,
                      degree_motherHS=0, degree_motherAA=0, degree_motherBA=0,
                      degree_motherG=0, degree_fatherHS=0, degree_fatherAA=0, 
                      degree_fatherBA=0, degree_fatherG=0,
                      foreign_born=0, foreign_born_father=0, foreign_born_mother=0,
                      income_trans=0, own_home=0, parents_married=0)
```
